---
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
header-includes:
  \usepackage{fontspec}
  \usepackage[utf8]{inputenc}
  \usepackage{pdfpages}
  \usepackage{graphicx}
  \setmainfont{Arial}
  \thispagestyle{empty}
---

```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(patchwork)
library(kableExtra)
library(rcartocolor)

palette = carto_pal(8,"Safe")

knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE)

theme_set(theme_minimal())

# source helper functions
source("../1-src/functions/calcRepea.R") # helper function for estimating precision/repeatability
source("../1-src/functions/regresAge.R") # helper function for calculating root mean squared error

```


```{r, echo=FALSE,message=FALSE, warning=FALSE, fig.width=8, fig.height=8}

comparison <- readRDS("../1-src/analysis/1-Hannum-clocks-Higgins/2-output/pheno_BloodRep_450K_Hannum_PC.Rds")

# a) original Hannum clock
fit <- lm(round(Hannum.EUTOPS, digits=4) ~ round(Hannum.higgins, digits=4), data = comparison)
anno <- paste0("R2 = ", signif(summary(fit)$r.squared, 3),
                 "\ny = ", round(fit$coefficients[2],2) , "x + ", round(fit$coefficients[1],2))

plota <- comparison %>%
  ggplot(aes(x = round(Hannum.higgins, digits=4), y = round(Hannum.EUTOPS, digits=4))) +
  geom_point() +
  theme(panel.background = element_blank(),aspect.ratio = 1) +
  geom_smooth(method = "lm",
                formula = y ~ x,
                size = 0.5,
                se = FALSE,
                alpha = 0.3,
                colour = "gray40") +
  xlab("Hannum age (higgins)") +
  ylab("Hannum age (EUTOPS)")  +
  xlim(35,85) +
  ylim(35,85) +
  annotate("text",
            x=40,
            y = 80,
            label = anno,
            hjust = 0,
            linewidth = 3)

# b) Hannum PC proxy

fit <- lm(round(Hannum_PCproxy.EUTOPS, digits=4) ~ round(Hannum_PCproxy.higgins, digits=4), data = comparison)
anno <- paste0("R2 = ", signif(summary(fit)$r.squared, 3),
                 "\ny = ", round(fit$coefficients[2],2) , "x + ", round(fit$coefficients[1],2))

plotb <- comparison %>%
  ggplot(aes(x = round(Hannum_PCproxy.higgins, digits=4), y = round(Hannum_PCproxy.EUTOPS, digits=4))) +
  geom_point() +
  theme(panel.background = element_blank(),aspect.ratio = 1) +
  geom_smooth(method = "lm, Hannum PC_prox",
                formula = y ~ x,
                size = 0.5,
                se = FALSE,
                alpha = 0.3,
                colour = "gray40") +
  xlab("Hannum_PCproxy age (higgins)") +
  ylab("Hannum_PCproxy age (EUTOPS)")  +
  xlim(35,85) +
  ylim(35,85) +
  annotate("text",
            x=40,
            y = 80,
            label = anno,
            hjust = 0,
            linewidth = 3)

### c) performance Hannum original and Hannum_PCy in 450K blood data ##----
result_BloodRep <- readRDS("../1-src/analysis/1-Hannum-clocks-Higgins/2-output/pheno_BloodRep_450K_Hannum_PC.Rds")

# create data frame for plotting
df <- as.data.frame(matrix(data=NA, nrow=2,ncol=2))
colnames(df) <- c("clock","RMSE")
clocks <- c("Hannum.EUTOPS","Hannum_PC.EUTOPS")

# accuracy for each clock, regress and save RMSE
for(i in 1:length(clocks)){
  df[i,1] <- clocks[i]
  df[i,2] <- sqrt(mean(regresAge(result_BloodRep,clocks[i],"", mode="fitting")$residuals^2)) 
}

# precision for each clock, calculate ICC
result_BloodRep$sample_ID <- result_BloodRep$sample
result_BloodRep$sample_type <- "Blood"
precision <- calcRepea(result_BloodRep,clocks)
precision <- precision[[2]] %>% dplyr::rename(clock=index) %>% select(-sample_type)

df <- full_join(df, precision)

#plot
cols <- c(palette[1],palette[5])
df$clock <- factor(df$clock, levels = clocks)
labels <- c("Hannum", "Hannum_PC")

plotc <- df %>%
  ggplot(aes(x=RMSE,y=ICC,color=clock, shape=clock)) +
  geom_point(size=4) +
  labs(color  = "", shape = "") +
  geom_errorbar(mapping=aes(ymin=CI0.95_lower, ymax=CI0.95_higher), width=0.2, size=0.2) + 
  xlim(3,4.5) +
  ylim(0.95,1) +
  scale_color_manual(labels=labels, values=cols)+
  scale_shape_manual(labels=labels, values=c(15,16))+
  theme_minimal()

# Original Hannum vs chronological age
fit <- lm(round(Hannum.EUTOPS, digits=4) ~ round(age, digits=4), data = result_BloodRep)
anno <- paste0("R2 = ", signif(summary(fit)$r.squared, 3),
               "\ny = ", round(fit$coefficients[2],2) , "x + ", round(fit$coefficients[1],2))

#CI for slope
#fit <- lm(Hannum.EUTOPS ~ age, data = result_BloodRep)
#confint(fit,'age',level=0.95)

plotd <- result_BloodRep %>%
  ggplot(aes(x = round(age, digits=4), y = round(Hannum.EUTOPS, digits=4))) +
  geom_point(color=cols[1], shape=15) +
  theme(panel.background = element_blank()) +
  geom_smooth(method = "lm",
              formula = y ~ x,
              size = 0.5,
              se = FALSE,
              alpha = 0.3,
              colour = "gray40") +
  xlab("Age") +
  ylab("Hannum predicted age")  +
  xlim(35,85) +
  ylim(35,85) +
  annotate("text",
           x=50,
           y = 40,
           label = anno,
           hjust = 0,
           size = 3) 

# PC Age vs chronological age
fit <- lm(round(Hannum_PC.EUTOPS, digits=4) ~ round(age, digits=4), data = result_BloodRep)
anno <- paste0("R2 = ", signif(summary(fit)$r.squared, 3),
               "\ny = ", round(fit$coefficients[2],2) , "x + ", round(fit$coefficients[1],2))

#CI for slope
#fit <- lm(Hannum_PC.EUTOPS ~ age, data = result_BloodRep)
#confint(fit,'age',level=0.95)
#fit <- lm(Hannum_PCproxy.EUTOPS ~ age, data = result_BloodRep)
#confint(fit,'age',level=0.95)

plote <- result_BloodRep %>%
  ggplot(aes(x = round(age, digits=4), y = round(Hannum_PC.EUTOPS, digits=4))) +
  geom_point(color=cols[2], shape=16) +
  theme(panel.background = element_blank(),plot.margin = margin(0, 0, 0, 0, "pt")) +
  geom_smooth(method = "lm",
              formula = y ~ x,
              size = 0.5,
              se = FALSE,
              alpha = 0.3,
              colour = "gray40") +
  xlab("Age") +
  ylab("Hannum_PC predicted age")  +
  xlim(35,85) +
  ylim(35,85) +
  annotate("text",
           x=50,
           y = 40,
           label = anno,
           hjust = 0,
           size = 3) 


## combined plot ##----

des <- "
#AABB
#AABB
cDDEE"

plot <- plota +  plotb + plotc + plotd + plote + plot_layout(design = des) + plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')

plot

```

\subsection{Extended data item 2. Reproducing clock calculations presented by [1].} Regression of age predictions for the same clock versions calculated ourselves (Y-axis) versus by [1] (X-axis) for **(a)** the original Hannum clock and **(b)** the PC proxy version of the Hannum clock. **(c)** Summary of the performance the clock predictions in the test set "BloodRep_450K". Precision (ICC; error bars represent 95% confidence intervals) and
accuracy (RMSE) of the linear regression between predicted age versus chronological
age, with corresponding regressions for **(d)** the original Hannum clock and **(e)** the PC version of the Hannum clock. The PC and PC proxy versions of the Hannum clock were trained to predict age and Hannum age, respectively, on the original subset of 78,464 CpGs by Higgins-Chen et al. [1] from the 450K "Hannum" dataset (n=656 samples; GEO ID GSE40279). Age estimations were done for 2x36 technical replicates from blood samples in the "BloodRep_450K" dataset (GEO ID GSE55763). <br>