---
title: "Modeling Accuracy (slope or MSE) and Precision for combined experiments"
output:
  html_document:
    code_folding: hide
    toc: false
    number_sections: false
---

##### Copyright (c) 2023 Charlotte D. Vavourakis. Licensed under the MIT license (see LICENSE.md).

```{r setup, include=FALSE, message=FALSE, echo=FALSE}


library(knitr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(kableExtra)

knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(echo=FALSE)

theme_set(theme_minimal())

dir.create("./4-output")
```

```{r, echo=FALSE, message=FALSE, echo=FALSE}

## combine results and save
experiment1 <- readRDS("./1-output/results_training_size.Rds") %>%
  mutate(modelf=case_when(method=="ElNet"&input.CpGs=="all" ~ "f1",
                          method=="ElNet"&input.CpGs=="unreliable.removed" ~ "f2",
                          method=="PC"&input.CpGs=="all" ~ "f3",
                          method=="PC"&input.CpGs=="unreliable.removed" ~ "f4",
                          method=="Lasso"&input.CpGs=="all" ~ "f5",
                          method=="Lasso"&input.CpGs=="unreliable.removed" ~ "f6"))

experiment2 <- readRDS("./2-output/results_training_size.Rds") %>%
  mutate(modelf=case_when(method=="ElNet"&input.CpGs=="all" ~ "f1",
                          method=="ElNet"&input.CpGs=="unreliable.removed" ~ "f2",
                          method=="PC"&input.CpGs=="all" ~ "f3",
                          method=="PC"&input.CpGs=="unreliable.removed" ~ "f4",
                          method=="Lasso"&input.CpGs=="all" ~ "f5",
                          method=="Lasso"&input.CpGs=="unreliable.removed" ~ "f6"))

experiment3 <- readRDS("./3-output/results_training_size.Rds") %>%
  mutate(modelf=case_when(method=="ElNet"&input.CpGs=="all" ~ "f1",
                          method=="ElNet"&input.CpGs=="unreliable.removed" ~ "f2",
                          method=="PC"&input.CpGs=="all" ~ "f3",
                          method=="PC"&input.CpGs=="unreliable.removed" ~ "f4",
                          method=="Lasso"&input.CpGs=="all" ~ "f5",
                          method=="Lasso"&input.CpGs=="unreliable.removed" ~ "f6"))

results_exp <- full_join(experiment1,experiment2)
results_exp <- full_join(results_exp,experiment3) %>% drop_na()

saveRDS(results_exp, file="./4-output/results_trainingsize_combined.Rds")




```



## Model accuracy

### slope regression

* modeled curves: <br>
     Y =  Ymax - aX^k ; <br>
     X =  number of training samples <br>
     Y = slope regression (age~predicted age in 3C external validation), Ymax is the maximum slope the model converges to <br>
     a,k = parameters determining curve shape <br>

```{r, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

# function
powerLaw1 <- function(x,a,b,c){
    c-(a*(x^b))
}

#### method 1
## model fits
# find starting parameters
ymax <- results_exp[results_exp$modelf == 'f1',] %>% pull(val.slope.regress) %>% max()
fit <- lm(log(val.slope.regress)~log(train.sample.size), data=results_exp[results_exp$modelf == 'f1',])
as <- unname(fit$coefficients[2]) #slope
bs <- unname(fit$coefficients[1]) #intercept
# check plot(fit)

test <- unique(results_exp$train.sample.size)
testy <- 1 - 100*(test**-1.2) #fiddled around with a
df <- data.frame(x=test,y=testy)
#plot(df)

check <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = 100, b = bs, c = ymax))

# fit and check fit
modelf1 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f1',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plota <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf1)[1], b = coef(modelf1)[2], c = coef(modelf1)[3]))

# save results
f1.fitted.values <- results_exp[results_exp$modelf == 'f1',] %>% mutate(val.slope.regress.fit = coef(modelf1)[3] - coef(modelf1)[1]*(train.sample.size**coef(modelf1)[2]))

doublecheck <- f1.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

#### method 2
# try same start parameters first
modelf2 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f2',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotb <- results_exp[results_exp$modelf == 'f2',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf2)[1], b = coef(modelf2)[2], c = coef(modelf2)[3]))

# save results
f2.fitted.values <- results_exp[results_exp$modelf == 'f2',] %>% mutate(val.slope.regress.fit = coef(modelf2)[3] - coef(modelf2)[1]*(train.sample.size**coef(modelf2)[2]))

doublecheck <- f2.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

results_exp_accuracy <- full_join(f1.fitted.values,f2.fitted.values)

#### method 3
# try same start parameters first
modelf3 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f3',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotc <- results_exp[results_exp$modelf == 'f3',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf3)[1], b = coef(modelf3)[2], c = coef(modelf3)[3]))

# save results
f3.fitted.values <- results_exp[results_exp$modelf == 'f3',] %>% mutate(val.slope.regress.fit = coef(modelf3)[3] - coef(modelf3)[1]*(train.sample.size**coef(modelf3)[2]))

doublecheck <- f3.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

results_exp_accuracy <- full_join(results_exp_accuracy,f3.fitted.values)

#### method 4
# try same start parameters first
modelf4 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f4',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotd <- results_exp[results_exp$modelf == 'f4',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf4)[1], b = coef(modelf4)[2], c = coef(modelf4)[3]))

# save results
f4.fitted.values <- results_exp[results_exp$modelf == 'f4',] %>% mutate(val.slope.regress.fit = coef(modelf4)[3] - coef(modelf4)[1]*(train.sample.size**coef(modelf4)[2]))

doublecheck <- f4.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

results_exp_accuracy <- full_join(results_exp_accuracy,f4.fitted.values)

#### method 5
# try same start parameters first
modelf5 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f5',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plote <- results_exp[results_exp$modelf == 'f5',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf5)[1], b = coef(modelf5)[2], c = coef(modelf5)[3]))

# save results
f5.fitted.values <- results_exp[results_exp$modelf == 'f5',] %>% mutate(val.slope.regress.fit = coef(modelf5)[3] - coef(modelf5)[1]*(train.sample.size**coef(modelf5)[2]))

doublecheck <- f5.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

results_exp_accuracy <- full_join(results_exp_accuracy,f5.fitted.values)

#### method 6
# try same start parameters first
modelf6 <- nls(val.slope.regress ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f6',], start = list(a=100, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotf <- results_exp[results_exp$modelf == 'f6',] %>%
  ggplot(aes(x=train.sample.size,y=val.slope.regress,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf6)[1], b = coef(modelf6)[2], c = coef(modelf6)[3]))

# save results
f6.fitted.values <- results_exp[results_exp$modelf == 'f6',] %>% mutate(val.slope.regress.fit = coef(modelf6)[3] - coef(modelf6)[1]*(train.sample.size**coef(modelf6)[2]))

doublecheck <- f6.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.slope.regress)) +
  geom_line(aes(y=val.slope.regress.fit))

results_exp_accuracy <- full_join(results_exp_accuracy,f6.fitted.values)

saveRDS(results_exp_accuracy, file = "./4-output/accuracy_slope.Rds")

#### plot all results and modeled curves
plot <- results_exp_accuracy %>%
  ggplot(aes(x=train.sample.size,color=input.CpGs)) +
  geom_point(aes(y=val.slope.regress), size=0.5) +
  geom_line(aes(y=val.slope.regress.fit), size=1.2) +
  facet_wrap(~method)

plot

#### save model paramaters
df <- data.frame(modelf=rep(NA,6), 
                 slope.a = rep(NA,6), slope.k = rep(NA,6), max.slope = rep(NA,6), 
                 SE.slope.a = rep(NA,6), SE.slope.k = rep(NA,6), SE.max.slope = rep(NA,6),
                 RSE= rep(NA,6), DF = rep(NA,6))
df[1,] <- c("f1",coef(modelf1),unname(summary(modelf1)$parameters[,2]), summary(modelf1)$sigma, summary(modelf1)$df[2])
df[2,] <- c("f2",coef(modelf2),unname(summary(modelf2)$parameters[,2]), summary(modelf2)$sigma, summary(modelf2)$df[2])
df[3,] <- c("f3",coef(modelf3),unname(summary(modelf3)$parameters[,2]), summary(modelf3)$sigma, summary(modelf3)$df[2])
df[4,] <- c("f4",coef(modelf4),unname(summary(modelf4)$parameters[,2]), summary(modelf4)$sigma, summary(modelf4)$df[2])
df[5,] <- c("f5",coef(modelf5),unname(summary(modelf5)$parameters[,2]), summary(modelf5)$sigma, summary(modelf5)$df[2])
df[6,] <- c("f6",coef(modelf6),unname(summary(modelf6)$parameters[,2]), summary(modelf6)$sigma, summary(modelf6)$df[2])

parTable <- df %>%
  mutate(method = case_when(
    modelf=="f1" ~ "ElNet", modelf=="f2" ~ "ElNet.unreliable.removed",
    modelf=="f3" ~ "PC.ElNet", modelf=="f4" ~ "PC.ElNet.unreliable.removed",
    modelf=="f5" ~ "Lasso", modelf=="f6" ~ "Lasso.unreliable.removed"
  )) %>%
  select(method,slope.a,SE.slope.a,slope.k,SE.slope.k,max.slope,SE.max.slope,RSE,DF)
parTable$slope.a <- round(as.numeric(parTable$slope.a), digits=2)
parTable$SE.slope.a <- round(as.numeric(parTable$SE.slope.a), digits=2)
parTable$slope.k <- round(as.numeric(parTable$slope.k), digits=2)
parTable$SE.slope.k <- round(as.numeric(parTable$SE.slope.k), digits=2)
parTable$max.slope <- round(as.numeric(parTable$max.slope), digits=3)
parTable$SE.max.slope <- round(as.numeric(parTable$SE.slope.k), digits=3)
parTable$RSE <- round(as.numeric(parTable$RSE), digits=2)

saveRDS(parTable, file="./4-output/accuracy_slope_modelpar.Rds")

```

```{r}
parTable %>% kbl() %>% kable_styling()
```

<br>

### MSE regression

* Modeled curves: <br>
     Y =  Ymin - aX^k ; <br>
     X =  number of training samples <br>
     Y = MSE regression (age~predicted age in 3C external validation), Ymin is the minimum MSE the model converges to <br>
     a,k = parameters determining curve shape <br>

```{r, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

#### method 1
## model fits

# find starting parameters
ymin <- results_exp[results_exp$modelf == 'f1',] %>% pull(val.MSE) %>% min()
fit <- lm(log(val.MSE)~log(train.sample.size), data=results_exp[results_exp$modelf == 'f1',])
as <- unname(fit$coefficients[2]) #slope
bs <- unname(fit$coefficients[1]) #intercept
# check plot(fit)

test <- unique(results_exp$train.sample.size)
testy1 <- ymin - (as*(test**-bs))
testy2 <- powerLaw1(x=test,a=as,b=-bs,c=ymin)
df1 <- data.frame(x=test,y=testy2, check="original")
testy2 <- powerLaw1(x=test,a=-10000,b=-1,c=ymin)
df2 <- data.frame(x=test,y=testy2, check="fiddling")
df <- full_join(df1,df2)
check <- df %>% ggplot(aes(x=x,y=y,color=check)) + geom_point()

check <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = -10000, b = -1, c = ymin))

# fit and check fit
modelf1 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f1',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plota <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf1)[1], b = coef(modelf1)[2], c = coef(modelf1)[3]))

# save results
f1.fitted.values <- results_exp[results_exp$modelf == 'f1',] %>% mutate(val.MSE.fit = coef(modelf1)[3] - coef(modelf1)[1]*(train.sample.size**coef(modelf1)[2]))

doublecheck <- f1.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

#### method 2
# try same start parameters first
modelf2 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f2',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plotb <- results_exp[results_exp$modelf == 'f2',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf2)[1], b = coef(modelf2)[2], c = coef(modelf2)[3]))

# save results
f2.fitted.values <- results_exp[results_exp$modelf == 'f2',] %>% mutate(val.MSE.fit = coef(modelf2)[3] - coef(modelf2)[1]*(train.sample.size**coef(modelf2)[2]))

doublecheck <- f2.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

results_exp_accuracy2 <- full_join(f1.fitted.values,f2.fitted.values)

#### method 3
# try same start parameters first
modelf3 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f3',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plotc <- results_exp[results_exp$modelf == 'f3',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf3)[1], b = coef(modelf3)[2], c = coef(modelf3)[3]))

# save results
f3.fitted.values <- results_exp[results_exp$modelf == 'f3',] %>% mutate(val.MSE.fit = coef(modelf3)[3] - coef(modelf3)[1]*(train.sample.size**coef(modelf3)[2]))

doublecheck <- f3.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

results_exp_accuracy2 <- full_join(results_exp_accuracy2,f3.fitted.values)

#### method 4
# try same start parameters first
modelf4 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f4',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plotd <- results_exp[results_exp$modelf == 'f4',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf4)[1], b = coef(modelf4)[2], c = coef(modelf4)[3]))

# save results
f4.fitted.values <- results_exp[results_exp$modelf == 'f4',] %>% mutate(val.MSE.fit = coef(modelf4)[3] - coef(modelf4)[1]*(train.sample.size**coef(modelf4)[2]))

doublecheck <- f4.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

results_exp_accuracy2 <- full_join(results_exp_accuracy2,f4.fitted.values)

#### method 5
# try same start parameters first
modelf5 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f5',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plote <- results_exp[results_exp$modelf == 'f5',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf5)[1], b = coef(modelf5)[2], c = coef(modelf5)[3]))

# save results
f5.fitted.values <- results_exp[results_exp$modelf == 'f5',] %>% mutate(val.MSE.fit = coef(modelf5)[3] - coef(modelf5)[1]*(train.sample.size**coef(modelf5)[2]))

doublecheck <- f5.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

results_exp_accuracy2 <- full_join(results_exp_accuracy2,f5.fitted.values)

#### method 6
# try same start parameters first
modelf6 <- nls(val.MSE ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f6',], start = list(a=-10000, b=-1, c=ymin), nls.control(maxiter = 1000, warnOnly = TRUE))

plotf <- results_exp[results_exp$modelf == 'f6',] %>%
  ggplot(aes(x=train.sample.size,y=val.MSE,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf6)[1], b = coef(modelf6)[2], c = coef(modelf6)[3]))

# save results
f6.fitted.values <- results_exp[results_exp$modelf == 'f6',] %>% mutate(val.MSE.fit = coef(modelf6)[3] - coef(modelf6)[1]*(train.sample.size**coef(modelf6)[2]))

doublecheck <- f6.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=val.MSE)) +
  geom_line(aes(y=val.MSE.fit))

results_exp_accuracy2 <- full_join(results_exp_accuracy2,f6.fitted.values)

saveRDS(results_exp_accuracy2, file = "./4-output/accuracy_MSE.Rds")

#### plot all results and modeled curves
plot <- results_exp_accuracy2 %>%
  ggplot(aes(x=train.sample.size,color=input.CpGs)) +
  geom_point(aes(y=val.MSE), size=0.5) +
  geom_line(aes(y=val.MSE.fit), size=1.2) +
  facet_wrap(~method)

plot

```

<br><br>

## Model precision

* modeled curves: <br>
     Y =  Ymax - aX^k ; <br>
     X =  number of training samples <br>
     Y = ICC in cervical smears repeatability set, Ymax is the maximum ICC the model converges to <br>
     a,k = parameters determining curve shape <br>
     
* Not best model fits for ElNet/Lasso, a lot of sample sets for which precision was low also for median training set sample size.

```{r, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

#### method 1
## model fits

# find starting parameters
ymax <- results_exp[results_exp$modelf == 'f1',] %>% pull(rep.ICC) %>% max()
fit <- lm(log(rep.ICC)~log(train.sample.size), data=results_exp[results_exp$modelf == 'f1',])
as <- unname(fit$coefficients[2]) #slope
bs <- unname(fit$coefficients[1]) #intercept

test <- unique(results_exp$train.sample.size)
testy <- ymax - as*(test**bs)
df <- data.frame(x=test,y=testy)
#plot(df)

check <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = 0.1, b = bs, c = ymax))

# fit and check fit
modelf1 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f1',], start = list(a=0.05, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plota <- results_exp[results_exp$modelf == 'f1',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf1)[1], b = coef(modelf1)[2], c = coef(modelf1)[3]))

# save results
f1.fitted.values <- results_exp[results_exp$modelf == 'f1',] %>% mutate(rep.ICC.fit = coef(modelf1)[3] - coef(modelf1)[1]*(train.sample.size**coef(modelf1)[2]))

doublecheck <- f1.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

#### method 2
# try same start parameters first
modelf2 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f2',], start = list(a=0.05, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotb <- results_exp[results_exp$modelf == 'f2',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf2)[1], b = coef(modelf2)[2], c = coef(modelf2)[3]))

# save results
f2.fitted.values <- results_exp[results_exp$modelf == 'f2',] %>% mutate(rep.ICC.fit = coef(modelf2)[3] - coef(modelf2)[1]*(train.sample.size**coef(modelf2)[2]))

doublecheck <- f2.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

results_exp_precision <- full_join(f1.fitted.values,f2.fitted.values)

#### method 3
# find starting parameters
ymax <- results_exp[results_exp$modelf == 'f3',] %>% pull(rep.ICC) %>% max()
fit <- lm(log(rep.ICC)~log(train.sample.size), data=results_exp[results_exp$modelf == 'f3',])
as <- unname(fit$coefficients[2]) #slope
bs <- unname(fit$coefficients[1]) #intercept

test <- unique(results_exp$train.sample.size)
testy <- ymax - as*(test**bs)
df <- data.frame(x=test,y=testy)
#plot(df)

check <- results_exp[results_exp$modelf == 'f3',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = 1, b = -1, c = ymax))

modelf3 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f3',], start = list(a=1, b=-1, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotc <- results_exp[results_exp$modelf == 'f3',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf3)[1], b = coef(modelf3)[2], c = coef(modelf3)[3]))

# save results
f3.fitted.values <- results_exp[results_exp$modelf == 'f3',] %>% mutate(rep.ICC.fit = coef(modelf3)[3] - coef(modelf3)[1]*(train.sample.size**coef(modelf3)[2]))

doublecheck <- f3.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

results_exp_precision <- full_join(results_exp_precision,f3.fitted.values)

#### method 4
# try same start parameters first
modelf4 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f4',], start = list(a=1, b=-1, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotd <- results_exp[results_exp$modelf == 'f4',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf4)[1], b = coef(modelf4)[2], c = coef(modelf4)[3]))

# save results
f4.fitted.values <- results_exp[results_exp$modelf == 'f4',] %>% mutate(rep.ICC.fit = coef(modelf4)[3] - coef(modelf4)[1]*(train.sample.size**coef(modelf4)[2]))

doublecheck <- f4.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

results_exp_precision <- full_join(results_exp_precision,f4.fitted.values)

#### method 5
# find start parameters
# find starting parameters
ymax <- results_exp[results_exp$modelf == 'f5',] %>% pull(rep.ICC) %>% max()
fit <- lm(log(rep.ICC)~log(train.sample.size), data=results_exp[results_exp$modelf == 'f5',])
as <- unname(fit$coefficients[2]) #slope
bs <- unname(fit$coefficients[1]) #intercept

modelf5 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f5',], start = list(a=0.1, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plote <- results_exp[results_exp$modelf == 'f5',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf5)[1], b = coef(modelf5)[2], c = coef(modelf5)[3]))

# save results
f5.fitted.values <- results_exp[results_exp$modelf == 'f5',] %>% mutate(rep.ICC.fit = coef(modelf5)[3] - coef(modelf5)[1]*(train.sample.size**coef(modelf5)[2]))

doublecheck <- f5.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

results_exp_precision <- full_join(results_exp_precision,f5.fitted.values)

#### method 6
# try same start parameters first
modelf6 <- nls(rep.ICC ~ c - (a*train.sample.size^b), results_exp[results_exp$modelf == 'f6',], start = list(a=0.1, b=bs, c=ymax), nls.control(maxiter = 1000, warnOnly = TRUE))

plotf <- results_exp[results_exp$modelf == 'f6',] %>%
  ggplot(aes(x=train.sample.size,y=rep.ICC,color=input.CpGs)) +
  geom_point() +
  geom_function(fun = powerLaw1, args = list(a = coef(modelf6)[1], b = coef(modelf6)[2], c = coef(modelf6)[3]))

# save results
f6.fitted.values <- results_exp[results_exp$modelf == 'f6',] %>% mutate(rep.ICC.fit = coef(modelf6)[3] - coef(modelf6)[1]*(train.sample.size**coef(modelf6)[2]))

doublecheck <- f6.fitted.values %>%
  ggplot(aes(x=train.sample.size)) +
  geom_point(aes(y=rep.ICC)) +
  geom_line(aes(y=rep.ICC.fit))

results_exp_precision <- full_join(results_exp_precision,f6.fitted.values)

saveRDS(results_exp_precision, file = "./4-output/precision_ICC.Rds")

#### plot all results and modeled curves
plot <- results_exp_precision %>%
  ggplot(aes(x=train.sample.size,color=input.CpGs)) +
  geom_point(aes(y=rep.ICC), size=0.5) +
  geom_line(aes(y=rep.ICC.fit), size=1.2) +
  facet_wrap(~method)

plot

#### save model paramaters
df <- data.frame(modelf=rep(NA,6), 
                 ICC.a = rep(NA,6), ICC.k = rep(NA,6), max.ICC = rep(NA,6), 
                 SE.ICC.a = rep(NA,6), SE.ICC.k = rep(NA,6), SE.max.ICC = rep(NA,6),
                 RSE= rep(NA,6), DF = rep(NA,6))
df[1,] <- c("f1",coef(modelf1),unname(summary(modelf1)$parameters[,2]), summary(modelf1)$sigma, summary(modelf1)$df[2])
df[2,] <- c("f2",coef(modelf2),unname(summary(modelf2)$parameters[,2]), summary(modelf2)$sigma, summary(modelf2)$df[2])
df[3,] <- c("f3",coef(modelf3),unname(summary(modelf3)$parameters[,2]), summary(modelf3)$sigma, summary(modelf3)$df[2])
df[4,] <- c("f4",coef(modelf4),unname(summary(modelf4)$parameters[,2]), summary(modelf4)$sigma, summary(modelf4)$df[2])
df[5,] <- c("f5",coef(modelf5),unname(summary(modelf5)$parameters[,2]), summary(modelf5)$sigma, summary(modelf5)$df[2])
df[6,] <- c("f6",coef(modelf6),unname(summary(modelf6)$parameters[,2]), summary(modelf6)$sigma, summary(modelf6)$df[2])

parTable <- df %>%
  mutate(method = case_when(
    modelf=="f1" ~ "ElNet", modelf=="f2" ~ "ElNet.unreliable.removed",
    modelf=="f3" ~ "PC.ElNet", modelf=="f4" ~ "PC.ElNet.unreliable.removed",
    modelf=="f5" ~ "Lasso", modelf=="f6" ~ "Lasso.unreliable.removed"
  )) %>%
  select(method,ICC.a,SE.ICC.a,ICC.k,SE.ICC.k,max.ICC,SE.max.ICC,RSE,DF)
parTable$ICC.a <- round(as.numeric(parTable$ICC.a), digits=1)
parTable$SE.ICC.a <- round(as.numeric(parTable$SE.ICC.a), digits=1)
parTable$ICC.k <- round(as.numeric(parTable$ICC.k), digits=1)
parTable$SE.ICC.k <- round(as.numeric(parTable$SE.ICC.k), digits=1)
parTable$max.ICC <- round(as.numeric(parTable$max.ICC), digits=1)
parTable$SE.max.ICC <- round(as.numeric(parTable$SE.ICC.k), digits=1)
parTable$RSE <- round(as.numeric(parTable$RSE), digits=3)

saveRDS(parTable, file = "./4-output/precision_ICC_modelpar.Rds")

```

```{r}
parTable %>% kbl() %>% kable_styling()
```

<br><br>

##### Final CpGs or PC components in the trained models

```{r, message=FALSE, echo=FALSE, fig.width=12, fig.height=4}

plot <- results_exp %>%
  ggplot(aes(x=train.sample.size,y=n.features,color=input.CpGs)) +
  geom_point() +
  #geom_smooth(method="lm", se=FALSE) +
  facet_wrap(~method)

plot

```
